### 2009-05-25 doc/manuals/janitor/tex_concept/concept.tex
Use pregenerated graphics instead of fancy LaTeX graphics style
Series of updates to the Janitor documentation.

 The main language for the implementation of the functionality of the
 dynamic runtime environment functionality is Perl. And it is solely required (exceptions are the
 integration with the Grid Manager and the Web server) for the Janitor. In
 the pre-web-service implementation the Catalog remains a static web page.
 The Perl code is split into multiple modules as depicted in Figure~\ref{fig:janitorDependencies}. 
 The modules can be separated into two
 functional groups. One addresses the retrieval of information from
 the Catalog RDF file in the left major branch of the figure. The other
 addresses the process of fetching and installing the packages.

 \begin{landscape}
\begin{figure}[!h]
\vspace{4cm}
 \begin{center} 
    \resizebox{24cm}{!}{\input{images/dependencies}}
    \mycaption{Modules of the Janitor and their dependencies}{}
 \end{center} 
\vfill
 \label{fig:janitorDependencies} 
\end{figure}
 \end{landscape}

@@

\section{Runtime environments states}
A main motivation for the managed, manual initiation of dynamic RE
installation is the subsequent manual verification of the installed
packages -- prior to their use in production.
With an automation of the installation, the verification of
that process shall be performed externally to that process.  At this
time, only the automation of the installation has been implemented.
To reflect the progress the external verification has made, REs
are said to be in states. The current implementation lists
installable REs aside the installed REs in the grid information system,
in order to stimulate grid clients to submit packages. The here described
states will be represented to the clients in upcoming developments.

These states are specific for every CE and communicated
between the Janitor and the Execution Service. Table~\ref{tab:states} shows all
possible states, while Figure~\ref{fig:RE_states}
displays the transitions between the states that a Runtime Environment may
be in during its life time at a particular CE. \task{?at a particular CE?}

@@

Upon presentation of a package to a Catalog, a CE may
classify a package to be \texttt{INSTALLABLE} if all the dependencies
are installable or already \texttt{INSTALLED}. The installation can be
performed manually (\texttt{INSTALLING}/{\bf m}) or in an automated fashion
(\ldots/{\bf a}). Should the installation process return an error, then
the installation has \texttt{FAILED}. Once the installation succeeded,
the installed package is validated for its correctness. Should that
process fail, then the package's state it is said to be \texttt{BROKEN}.

@@

Janitor has two states for jobs: \texttt{PREPARED} and \texttt{INITIALIZED}. After a job has been succesfully registered 
in Janitor, its state will be set to \texttt{PREPARED}. Invalid jobs are not cached in Janitor. 
After Janitor is requested to deploy the runtime environment, the state of the job will change to \texttt{INITIALIZED}. If 
an unforeseen exception occures during that process, Janitor will drop the job from its database and set the affected runtime
environments to the state \texttt{FAILED}.

\section{Integration into AREX}

@@

Currently the Janitor perl scripts must be executed as root. In order to execute Janitor with the required permissions,
a setuid wrapper has been written. During the installation \task{not done yet} the mode of rjanitor will be set to \texttt{u+s}
to make the binary suid root. \task{According to Daniel: ``A future version of the Janitor will get rid of the suid root helper.''}\\

Thus, Janitor is not executed directly. Instead a dynamic link will be created in \task{libexec or sbin} which is pointing to
the wrapper.\\

@@

Security is a major concern for the grid systems. An automatic software
installation inheritently introduces security threats. This section
addresses those and describes the available solutions to limit
security risks.

In the current installation, every user authorised to execute a job is
also authorised to install a REs. Restrictions are
only imposed on the set of DRE that are available for installation.
Restrictions are imposed by the site admininistrators on the descriptions
that are given by the Catalog that is offering the package. These
descriptions may explicitly mention DRE names, a regular expression on
these, or refer to tags of packages that categorise these. However,
the core of these controls lie with the maintainers of the Catalog,
who needs to be trusted.

@@

\begin{itemize}
\item DREs are not distinguished between {\em installed} and
  {\em installable} in the information system.
\item No status information is given at the time an DRE is installed.
\end{itemize}

Malevolent regular users with respective training in using system exploits
to gain root access are likely to find security holes by regularly
submitted scripts.  The authentication and authentification of users,
together with respective logging, is the major defense against such
attacks. What is consequently left to be protected against are unwanted
sideeffects by the installation of software.

@@

The installation of packages from the Debian distribution (or other
packages of mainstream Linux distributions) is seeked to reduce the
complexity and burden in the maintenance for DRE. In the
current implementation Debian packages may be installed only by their
transformation into tar files. With the advent of the interface for
the virtualisation of the grid infrastructure, it is anticipated to
work with native packages of the Debian Linux distribution. The reuse
of packages that passed many eyeballs - as it is the case with packages
from major Linux distribution - security is further increased or becomes
as high as with the operating system underneath virtual clients.

>>>>

 The main language for the implementation of the functionality of the
 dynamic runtime environment functionality is Perl. And it is solely required (exceptions are the
 integration with the Grid Manager and the Web server) for the Janitor. In
 the pre-web-service implementation the Catalog remains a static web page.
 The Perl code is split into multiple modules as depicted in Figure~\ref{fig:janitorDependencies}. 
 The modules can be separated into two
 functional groups. One addresses the retrieval of information from
 the Catalog's RDF file in the left major branch of the figure. The other
 addresses the process of fetching and installing the packages.

 \begin{landscape}
\begin{figure}[!h]
\vspace{4cm}
 \begin{center} 
%    \resizebox{24cm}{!}{\input{images/dependencies}}
    \includegraphics[width=24cm]{images/dependencies.png}
    \mycaption{Modules of the Janitor and their dependencies}{}
 \end{center} 
\vfill
 \label{fig:janitorDependencies} 
\end{figure}
 \end{landscape}

@@

\section{States of runtime environments}
A major motivation for the managed, manual initiation of dynamic RE
installation is the subsequent manual verification of the installed
packages -- prior to their use in production.
With an automation of the installation, the verification of
that process shall be performed externally to that process.  At this
time, only the automation of the installation has been implemented.
To reflect the progress the external verification has made, REs
are said to be in states. The current implementation lists
installable REs aside the installed REs in the grid information system,
in order to stimulate grid clients to submit packages. The here described
states will be represented to the clients in upcoming developments.

These states are specific for every compute element (CE) and communicated
between the Janitor and the Execution Service. Table~\ref{tab:states} shows all
possible states, while Figure~\ref{fig:RE_states}
displays the transitions between the states that a Runtime Environment may
be in during its life time at a particular CE. \task{?at a particular CE?}

@@

Upon presentation of a the package name to a Catalog, from which details about
the package are retrieved, a CE may
classify a package to be \texttt{INSTALLABLE} if all the dependencies
are installable or already \texttt{INSTALLED}. The installation can be
performed manually (\texttt{INSTALLING}/{\bf m}) or in an automated fashion
(\ldots/{\bf a}). Should the installation process return an error, then
the installation has \texttt{FAILED}. Once the installation succeeded,
the installed package is validated for its correctness. Should that
process fail, then the package's state it is said to be \texttt{BROKEN}.

@@

The Janitor manages the states that the runtime environments at a particular
compute element are in. However, it is also most important for the Janitor
to be aware of the jobs that depend on the installtion of a RE. REs still in use
should not be removed until the respective job has completed its computations.
The installation or removal of REs by the Janitor is perceived as a mere
consequene of jobs demanding a RE or not, thus, the communcation between
the job-manager AREX and the Janitor will be performed on that 'job level'.

The Janitor has two states for jobs: \texttt{PREPARED} and \texttt{INITIALIZED}. After a job has been succesfully registered 
in the Janitor, its state will be set to \texttt{PREPARED}. Invalid jobs are not cached. 
After the Janitor is requested to deploy the runtime environment, the state of the job will change to \texttt{INITIALIZED}. If 
an unforeseen exception occures during that process, the Janitor will drop the job from its database and set the affected runtime
environments to the state \texttt{FAILED}.

\section{Integration with AREX}

@@

In the current implementation the Janitor perl scripts must be executed as root. In order to execute Janitor with the required permissions,
a setuid wrapper has been prepared. During the installation \task{not done yet} the mode of rjanitor will be set to \texttt{u+s},
i.e. the sticky bit is set to make the binary suid root. \task{According to Daniel: ``A future version of the Janitor will get rid of the suid root helper.''}\\

Thus, Janitor is not executed directly. Instead, a dynamic link will be created in \task{libexec or sbin} which is pointing to
the wrapper.\\

@@

Security is a major concern for grid systems. Any additional feature
and especially an automatic software
installation inheritently introduces security threats. This section
addresses those and describes the available solutions to limit
security risks.

In the current installation, every user authorised to execute a job is
also authorised to install a REs. Restrictions are
only imposed on the set of dynamic REs that are available for installation.
Restrictions are imposed by the site admininistrators on the descriptions
that are given by the Catalog that is offering the package. These
descriptions may explicitly mention dynamic REs' names, e.g. a regular expression on
these, or refer to tags of packages that categorise these. However,
the core of these controls lies with the maintainers of the Catalog,
who needs to be trusted.

@@

\begin{itemize}
\item Dynamic REs are not distinguished between {\em installed} and
  {\em installable} in the information system.
\item No status information is given at the time a dynamic RE is installed.
\end{itemize}

Malevolent regular users with respective training in using system exploits
to gain root access are likely to find security holes by regularly
submitted scripts.  The authentication and authentification of users,
together with respective logging, is the major defense against such
attacks. What is consequently left to be protected against are unwanted
side-effects by the installation of software.

@@

The installation of packages from the Debian distribution (or other
packages of mainstream Linux distributions) is seeked to reduce the
complexity and burden in the maintenance for dynamic REs. In the
current implementation, Debian packages may be installed only by their
transformation into tar files. With the advent of the interface for
the virtualisation of the grid infrastructure, it is anticipated to
work with native packages of the Debian Linux distribution. The reuse
of packages that passed many eyeballs - as it is the case with packages
from major Linux distribution - security is further increased or becomes
as high as with the operating system underneath virtual clients.

### 2009-05-25 doc/manuals/janitor/tex_installation/installation.tex
Series of updates to the Janitor documentation.

% ** Dependencies
% *** Ordianary: log4perl and ``Redland RDF Library - Perl Interface''
% *** WebService: libperl-dev
% *** Protege: For maintaining the knowledge base

In order to use Janitor, the two perl packages listed in table~\ref{tab:install_dependencies} are required. To have the WebService interface for Janitor, the packages listed in the table~\ref{tab:install_dependencies_optional} 
have to be installed before the build process is.
\begin{table}[!h]
   \begin{center}
        \mycaption{Required perl packages for Janitor.}{Log4perl is used for the internal logging of Janitor, while the 
                   Redland RDF libraray is used for accessing the knowledge base of RTEs.}
        \label{tab:install_dependencies}
	\begin{tabular}{|p{3cm}|p{7cm}|}
	\hline
	   liblog-log4perl-perl & \textit{Log4perl is a port of the log4j logging package}\\
	\hline
	   librdf-perl          & \textit{Perl language bindings for the Redland RDF library}\\
	\hline
	\end{tabular} 
   \end{center}
\end{table}
\begin{table}[!h]
   \begin{center}
        \mycaption{Optional libraries for Janitor.}{The library libperl-dev provides header files which are needed to link the 
                   WebService to the Perl interpreter.}
        \label{tab:install_dependencies_optional}
	\begin{tabular}{|p{3cm}|p{7cm}|}
	\hline
	   libperl-dev & \textit{ Perl library: development files}\\
	\hline
	\end{tabular}
   \end{center}
\end{table}
\forcelinebreak

Janitor will be installed by default along with ARC1. If desired it is possible to disable the shipping of Janitor 
using the \textit{configure} flags \textit{--disable-janitor-service} for the complete janitor or 
\textit{--disable-janitor-webservice} for only the Web Service support.
Furthermore it is recommended to install the ontology editor 
Prot\`eg\`e\footnote{\href{http://protege.stanford.edu}{http://protege.stanford.edu}} in order to be to easily maintain the 
knowledge database of installable packages.

@@

The current version of Janitor can be configured using the common file \textit{arc.conf} which is to be found in the 
configuration directory \textit{etc}. Janitor is using the environment variable \textit{NORDUGRID\_CONFIG} to determine the location 
of the corresponding file. If the variable is not set, the default location \textit{/etc/arc.conf} will be used. 
The configuration is assigned by the section \lbrack janitor\rbrack. The table~\ref{tab:arc_conf_tags} contains the available tags
for janitor configuration.
\begin{landscape}
\begin{table}[!h]
   \begin{center}
        \mycaption{Tags usable in \textit{arc.conf} within the section janitor.}{Tags usable in \textit{arc.conf} within the section janitor.}
	\label{tab:arc_conf_tags}
	\begin{longtable}{|p{3cm}|p{10cm}||p{10cm}|}
	\hline
	   \textbf{tag}    & \textbf{example}                      & \textbf{description}\\
        \hline
           enabled         & "1"                                   & Boolean flag which enables or disables janitor in A-REX.\\
	   uid             & "root"                                & The effective uid. \\
	   gid             & "0"                                   & The effective gid. \\
	   registrationdir & "/var/spool/nordugrid/janitor"        & Directory where we the current states of jobs are kept. \\
	   catalog         & "/var/spool/nordugrid/janitor/catalog/knowarc.rdf"& URL of the catalog containg the package information.\\
	   downloaddir     & "/var/spool/nordugrid/janitor/download" & Directory for downloads \\
	   installationdir & "/var/spool/nordugrid/janitor/runtime"& Directory for installation of packages                   \\
	   jobexpirytime   & "7200"                                & If a job is older than this, it is considered dead and assigned to be removal pending.\\
	   rteexpirytime   & "36"                                  & If a runtime environment was not used for this time, it will be assigned to be removal pending.\\
	   allow\_base     & "*"                                   & Allow rule for base packages. \\
	   deny\_base      & "debian::etch"                        & Deny rule for base packages.\\
	   allow\_rte      & "*"                                   & Allow rule for base packages. \\
	   deny\_rte       & "APPS/MATH/ELMER-5.0.2"               & Deny rule for base packages. \\
	   logconf         & "/opt/nordugrid/etc/log.conf"         & Location of the logging configuration file for janitor.\\
	\hline
	\end{longtable}
   \end{center}
\end{table}
\end{landscape}

The parameter \texttt{enable} defines whether Janitor shall be used within A-REX or not. Use the value \texttt{"0"} to disable Janitor. The
\texttt{uid} and the \texttt{gid} are defining which effective the user and group id shall be used for Janitor. 
The \texttt{registrationdir} describes the directory in which the subdirectories \texttt{jobs} and \texttt{rtes} will be created.
In these directories the states of the jobs and the runtime environments will be stored. 
The knowledge base of installable packages is specified by the parmeter \texttt{catalog}. 
Its value can be any kind of URL pointing to an file written in the Resource Description Framework (RDF) format.
The specification of the RDF file will be explained in detail in section~\ref{sec:catalog}.
The parameter \texttt{downloaddir} assignes the directory in which the installation files will be saved after they have been 
downloaded or copied from the repository which was specified by the catalog. The \texttt{installationdir} finally specifies the
directory into which all packages will be installed. Instead of the other directories the \texttt{installationdir} should be 
available for all computing elements i.e. by using a shared volume.
If the configuration file furthermore contains the \texttt{runtimedir} tag within the section \texttt{grid-manager}, Janitor will also
create a symbolic link in the \texttt{runtimedir} pointing to the configuration script of the installation done by Janitor.
The tags \texttt{jobexpirytime} and \texttt{rteexpirytime} are used for automated cleanup and is defined in seconds.
The default value for the \texttt{jobexpirytime} is seven days and for the \texttt{rteexpirytime} three days. 
The additional tags \texttt{allow\_base} \texttt{deny\_base} \texttt{allow\_rte} and \texttt{deny\_rte} are used to include
or exclude certain base packages or runtime environments of the catalog. This feature is useful, if the catalog is maintain by
a higher organization.
The path to the log4perl configuration file is defined by the tag \texttt{logconf}.
An examples how to configure arc and log4perl is provided in the Listings~\ref{lst:arc_conf} and ~\ref{lst:log_conf} and.

@@

Janitor is designed for the usage on Linux distributions.

>>>>

% ** Dependencies
% *** Ordinary: log4perl and ``Redland RDF Library - Perl Interface''
% *** WebService: libperl-dev
% *** Protege: For maintaining the knowledge base

In order to use the Janitor, the two perl packages listed in
table~\ref{tab:install_dependencies} are required. To have
the WebService interface for the Janitor, the packages listed in the
table~\ref{tab:install_dependencies_optional} need to be installed before
the build process is.

\begin{table}[!h]
   \begin{center}
        \mycaption{Required perl packages for Janitor.}{Log4perl is used for the internal logging of Janitor, while the 
                   Redland RDF libraray is used for accessing the knowledge base of Runtime Environments.}
        \label{tab:install_dependencies}
	\begin{tabular}{|p{3cm}|p{7cm}|}
	\hline
	   liblog-log4perl-perl & \textit{Log4perl is a port of the log4j logging package}\\
	\hline
	   librdf-perl          & \textit{Perl language bindings for the Redland RDF library}\\
	\hline
	\end{tabular} 
   \end{center}
\end{table}
\begin{table}[!h]
   \begin{center}
        \mycaption{Optional libraries for Janitor.}{The library libperl-dev provides header files which are needed to link the 
                   WebService to the Perl interpreter.}
        \label{tab:install_dependencies_optional}
	\begin{tabular}{|p{3cm}|p{7cm}|}
	\hline
	   libperl-dev & \textit{ Perl library: development files}\\
	\hline
	\end{tabular}
   \end{center}
\end{table}
\forcelinebreak

The Janitor will be installed by default along with ARC1. Support for
the dynamic REs can be completely disabled at compile time by using
the \textit{configure} flags \textit{--disable-janitor-service} for
the complete janitor or \textit{--disable-janitor-webservice} to only
disable the Janitor's Web Service.

Furthermore it is recommended to install the ontology editor 
Prot\`eg\`e\footnote{\href{http://protege.stanford.edu}{http://protege.stanford.edu}} in order to be to easily maintain the 
knowledge database of installable packages.

@@

The current version of the Janitor can be configured using the common file \textit{arc.conf}. It is expected in the 
configuration directory \textit{etc}. The Janitor is using the environment variable \textit{NORDUGRID\_CONFIG}
to determine the location of the corresponding file. If that variable is not set, the default location \textit{/etc/arc.conf} will be used. 
The configuration is assigned by the section \lbrack janitor\rbrack. The table~\ref{tab:arc_conf_tags} contains the available tags
for janitor configuration.

\begin{landscape}
\begin{table}[!h]
   \begin{center}
        \mycaption{Tags usable in \textit{arc.conf} within the section janitor.}{Tags usable in \textit{arc.conf} within the section janitor.}
	\label{tab:arc_conf_tags}
	\begin{longtable}{|p{3cm}|p{10cm}||p{10cm}|}
	\hline
	   \textbf{tag}    & \textbf{example}                      & \textbf{description}\\
        \hline
           enabled         & "1"                                   & Boolean flag which enables or disables janitor in A-REX.\\
	   uid             & "root"                                & The effective uid. \\
	   gid             & "0"                                   & The effective gid. \\
	   registrationdir & "/var/spool/nordugrid/janitor"        & Directory where we the current states of jobs are kept. \\
	   catalog         & "/var/spool/nordugrid/janitor/catalog/knowarc.rdf"& URL of the catalog containg the package information.\\
	   downloaddir     & "/var/spool/nordugrid/janitor/download" & Directory for downloads \\
	   installationdir & "/var/spool/nordugrid/janitor/runtime"& Directory for installation of packages                   \\
	   jobexpirytime   & "7200"                                & If a job is older than this, it is considered dead and assigned to be removal pending.\\
	   rteexpirytime   & "36"                                  & If a runtime environment was not used for this time, it will be assigned to be removal pending.\\
	   allow\_base     & "*"                                   & Allow rule for base packages. \\
	   deny\_base      & "debian::etch"                        & Deny rule for base packages.\\
	   allow\_rte      & "*"                                   & Allow rule for base packages. \\
	   deny\_rte       & "APPS/MATH/ELMER-5.0.2"               & Deny rule for base packages. \\
	   logconf         & "/opt/nordugrid/etc/log.conf"         & Location of the logging configuration file for janitor.\\
	\hline
	\end{longtable}
   \end{center}
\end{table}
\end{landscape}

The parameter \texttt{enable} defines whether the Janitor shall be used within A-REX or not. Use the value \texttt{"0"} to disable Janitor. The
\texttt{uid} and the \texttt{gid} are defining which effective the user and group id shall be used for Janitor. 
The \texttt{registrationdir} describes the directory in which the subdirectories \texttt{jobs} and \texttt{rtes} will be created.
In these directories the states of the jobs and the runtime environments will be stored. 
The knowledge base of installable packages is specified by the parameter \texttt{catalog}. 
Its value can be any kind of URL pointing to a file written in the Resource Description Framework (RDF) format.
The specification of the RDF file will be explained in detail in section~\ref{sec:catalog}.
The parameter \texttt{downloaddir} assigns the directory in which the installation files will be saved after they have been 
downloaded or copied from the repository which was specified by the catalog. The \texttt{installationdir} finally specifies the
directory into which all packages will be installed. Instead of the other directories, the \texttt{installationdir} should be 
available for all computing elements, i.e. by using a shared volume.
If the configuration file furthermore contains the \texttt{runtimedir} tag within the section \texttt{grid-manager}, the Janitor will also
create a symbolic link in the \texttt{runtimedir} pointing to the configuration script of the installation done by the Janitor.
The tags \texttt{jobexpirytime} and \texttt{rteexpirytime} are used for automated cleanup and is defined in seconds.
The default value for the \texttt{jobexpirytime} is seven days and for the \texttt{rteexpirytime} three days. 
The additional tags \texttt{allow\_base} \texttt{deny\_base} \texttt{allow\_rte} and \texttt{deny\_rte} are used to include
or exclude certain base packages or runtime environments of the catalog. This feature is useful, if the catalog is maintained by
a higher organization.
The path to the log4perl configuration file is defined by the tag \texttt{logconf}.
An examples how to configure arc and log4perl is provided in the Listings~\ref{lst:arc_conf} and ~\ref{lst:log_conf} and.

@@

The Janitor is designed for the usage on Linux distributions.
Its porting to other platforms has not yet been addressed.

The ARC middleware is not ultimately essential for dynamic Runtime Environments.
All the Perl code would be functional with any Grid middleware.

### 2009-05-25 doc/manuals/janitor/tex_titlepage/titlepage.tex
Series of updates to the Janitor documentation.

\begin{tabular}{rl}
\resizebox*{3cm}{!}{\includegraphics{images/ng-logo.png}}
&\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
\end{tabular}

@@

{\raggedleft NORDUGRID-MANUAL-9\par}

@@

%%%%---- The title ----
{\centering \textsc{\Large Web Service programming Tutorial}
\\\vspace{1cm} \normalsize\textcolor{discreeturgent}{This document is still under development, but don't hesitate sending your comments and suggestions to glodek@inb.uni-luebeck.de.} % DELETE ONLY THAT LINE!!!
\Large \par}
\vspace*{0.5cm}

@@

\vspace*{1.5cm}
%%%%---- A list of authors ----
    {\centering \large Michael Glodek\footnote{glodek@inb.uni-luebeck.de}, Steffen M\"oller \par}

>>>>

\begin{tabular}{rl}
\resizebox*{3cm}{!}{\includegraphics{ng-logo.png}}
&\parbox[b]{2cm}{\textbf \it {\hspace*{-1.5cm}NORDUGRID\vspace*{0.5cm}}}
\end{tabular}

@@

{\raggedleft NORDUGRID-D2.5-2 ???\par}

@@

%%%%---- The title ----
{\centering \textsc{\Large Dynamic Runtime Environment Installation with Janitor}
\\\vspace{1cm} \normalsize\textcolor{discreeturgent}{This document is still under development, but don't hesitate sending your comments and suggestions to glodek@inb.uni-luebeck.de.} % DELETE ONLY THAT LINE!!!
\Large \par}
\vspace*{0.5cm}

@@

\vspace*{1.5cm}
%%%%---- A list of authors ----
    {\centering \large Michael Glodek\footnote{glodek@inb.uni-luebeck.de}, Daniel Bayer,  Steffen M\"oller\footnote{moeller@inb.uni-luebeck.de} \par}
